Some thought experiments on the applicability of Granger causality and Directed Information in inferring direction of information flows

### 2015-07-06 paper.tex
Modify abstract.

\begin{abstract}
%\boldmath
Not without controversy, Granger causality and, more recently, directed information, are two measures used to estimate ``causal influence'' of one stochastic process on another. Recently, many works interpret direction of causal influence as the direction of ``information flow.'' To test this latter interpretation, this paper uses two simple experiments to test whether these measures predict the correct direction of information flow, \textit{i.e.}, from the transmitting to the receiving node. These measures have been used extensively in neuroscience, and complex interconnected systems in general, to infer causal influences, and more recently, to infer direction of information flows. Using the problem of communication across feedback channels (first with noiseless feedback, and then with noisy feedback), and using a strategy inspired by the work of Schalkwijk and Kailath, we show that the direction of information flow can be opposite to that inferred using Granger causality and Directed Information. Thus, while it might be reasonable to infer direction of causal influence using these techniques, one needs to exercise care in interpreting the direction of causal influence as the direction of information flow. 
\end{abstract}
\IEEEpeerreviewmaketitle

>>>>

% Abstract --------------------------------------------------------------------

\begin{abstract}
	Granger causality is an established measure of the ``causal influence'' that one stochastic process has on another. Along with its more recent generalization -- Directed Information -- Granger Causality has been used extensively in neuroscience, and in complex interconnected systems in general, to infer causal influences.     Of late, many works have begun to interpret the direction of causal influence as the direction of ``information flow'' {\color{red}(this is akin to, but different from, a criticism of Granger Causality that compares it to true causality, as defined by Judea Pearl)}. We test whether these measures correctly predict the direction of information flow by using two simple theoretical experiments, in which the true direction of information flow is known by design. The experiments are based on a communication system with a feedback channel, upon which a strategy inspired by the work of Schalkwijk and Kailath is employed. We show that in these experiments, the direction of information flow can be opposite to that which is inferred by using Granger causality and Directed Information. Thus, while it might be reasonable to infer the direction of causal influence using these techniques, one needs to exercise care in interpreting the direction of causal influence as the direction of information flow.
\end{abstract}
\IEEEpeerreviewmaketitle
